---
title: "Solutions for Lab 1"
author: "Leo lahti & Rajesh Shigdel"
date: "`r Sys.Date()`"
output: html_document
---



#command to load that Rds file (  )

# Tutorial Aims 
In this tutorial we will learn feature selection, dimension reduction, clustering, visualization, analysis and interpretation  using  
 miaverse (mia = MIcrobiome Analysis) and miaViz - (Microbiome Analysis Plotting and Visualization) to explore patterns in human gut microbiome datasets. 

The example in this tutorial are mainly based on this online book        
[Orchestrating Microbiome Analysis](https://microbiome.github.io/OMA/)

In this tutorial we use data from [The Human Gut Microbiome Atlas (HGMA)](https://www.microbiomeatlas.org/)



# Lab 1: feature selection & dimension reduction

## Load the data 

Let us load the readily processed data (TreeSummarizedExperiment object).

You can download the R data file tse.Rds from [here](https://github.com/microbiome/course_2021_ml4microbiome/blob/master/tse.Rds).

```{r}
tse <- readRDS("tse.Rds")
```

If you like to see how the data was created, you can download the
[zip](data_zipfile.zip) and run the data preparation
[script](tse_script.R).


## Load packages in R

```{r}
library(mia)
library(miaViz)
```

## Tasks


1. Aggregate the data to Phylum level 

```{r}
tse_phylum <- agglomerateByRank(tse, rank = "phylum")
```



2. Subset the data at the Species level with only taking bacterial Species that are greater than 10% of prevalence in the total sample


```{r}
altExp(tse,"species") <- agglomerateByRank(tse, "species")
getPrevalentTaxa(tse, detection = 0, prevalence = 10/100)
prev <- getPrevalentTaxa(tse, detection = 0, prevalence = 10/100)
                         
```


3. Subset a tse object which consist only two phyla i.e "Actinobacteria", "Cyanobacteria"


```{r}
tse_subset_by_feature <- tse[rowData(tse)$phylum %in% c("Actinobacteria", "Cyanobacteria")]
tse_subset_by_feature
```

 
4. Calculate relative abundances, and store the table to assays


```{r}
tse <- relAbundanceCounts(tse)

```


5.Perform a centered log-ratio transformation (clr) (i.e. mia::transformSamples)

```{r}
tse <- transformSamples(x = tse, abund_values = "counts", method = "clr", 
                        pseudocount = 1, name = "CLR")
```